---
title: "R Notebook"
output: html_notebook
---
 

```{r}
library(readxl)

# load and save data as df
df <- read_excel("LungCapData2.xlsx") # insert data path
# View first 10 entries of the data
head(df,10)


```

```{r}
# view dimension of data
dim(df)
```

```{r}
# view class type for each variable in data 
sapply(df, class)
```



```{r}
# view structure of data
str(df)

```

```{r}
# get summary statistics of data
summary(df)
```

```{r}

# calculate standard deviation for all attributes
sapply(df[,1:3], sd)
```


```{r}
library(e1071)
# check skewness of the data

# calculate skewness for each variable
skew <- apply(df[,1:3], 2, skewness)
print(skew)


#The further the distribution of the skew value from zero, the larger the skew to the left (negative skew value) or right (positive skew value).
```


```{r}
# view the correlation between for the numeric variables

correlations <- cor(df[,1:3])
print(correlations)
```



```{r}
#define variables

y <- df$LungCap
x <- df$Height
Xsquared <- x^2
Xcube <- x^3

```



```{r}
# visualize LungCap vs Height
plot(x,y, pch = 20, main = "Lung Capacity Vs Height", 
     ylab = "Lung Capacity",
     xlab = "Height")

```


```{r}
# formulate models
model1 = lm(y ~ x)
```

```{r}
summary(model1)
```

```{r}
par(mfrow=c(2,2))
plot(model1)
```


```{r}
model2 = lm(y ~ x + Xsquared)
summary(model2)
```

```{r}
par(mfrow=c(2,2))
plot(model2)

```


```{r}
model3 = lm(y ~ x + Xsquared + Xcube)
summary(model3)
```

```{r}
par(mfrow=c(2,2))
plot(model3)
```



```{r}
# define variable of x (height) values to fit  model2 and model3

Xvalues<- seq(min(x),max(x),0.1)
ypred_2 <-predict(model2, list(x=Xvalues, Xsquared = Xvalues^2)) # for model2
ypred_3 <-predict(model3, list(x=Xvalues, Xsquared = Xvalues^2, Xcube = Xvalues^3)) # for model3
```




```{r}
#fit model into data
plot(x,y, pch =20, main = "Lung Capacity Vs Height", 
     ylab = "Lung Capacity",
     xlab = "Height")
abline(model1, lwd=2,col = "green")
lines(Xvalues,ypred_2, lwd=2, col = "red")
lines(Xvalues,ypred_3, lwd=2,col = "blue")


```

```{r}

#show plots serparately since model2 and model3 are overlapping

#model1
par(mfrow=c(2,2))
plot(x,y, pch =20, main = "Model One", 
     ylab = "Lung Capacity",
     xlab = "Height")
abline(model1, lwd=2,col = "green")

#model2
plot(x,y, pch =20, main = "Model Two", 
     ylab = "Lung Capacity",
     xlab = "Height")
lines(Xvalues,ypred_2, lwd=2, col = "red")

#model3
plot(x,y, pch =20, main = "Model Three", 
     ylab = "Lung Capacity",
     xlab = "Height")
lines(Xvalues,ypred_3, lwd=2,col = "blue")
```


```{r}
# Anova comparation of the 3 models
anova(model1,model2,model3)
```



```{r}
# Anova comparation of model1 and model2
anova(model1, model2)
```

```{r}
# Anova comparation of model2 and model3
anova(model2,model3)
```



Conclusion

Regarding the results from the above ANOVA Analysis of the models, The p-value comparing Model 1 to the quadratic Model 2 lacks a numerical value for model 1, which could be considered as zero value and also an indication that a linear fit is not sufficient. However, the p-value comparing the quadratic Model 2 to the cubic Model 3 is less than 0.05, which is very significant. Consequently, the quadratic fit is very sufficient. Model 3 seems unnecessary because its p-value is 0.652306,  Hence, the quadratic function appear to provide a reasonable fit to the data. In addition, a large f value (58.83337004)for model 2 depicts a significant variation among the data, variation among the height and lung capacity.

